Should geophysicists use gravity anomaly or gravity disturbance?

### 2018-05-18 manuscript/letter.tex
The Bouguer correction part deserves a section
This is different enough from the discussion about the disturbance vs
free-air to deserve its own section. Plus, this way we end the previous
section with the part about the magnitude of the error from using the
anomaly.
Reworked the part about Bouguer and its meaning
This part was a bit disjointed and it wasn't address the important
question about the meaning of the corrections. In applied geophysics,
Bouguer is more important that free-air and more attention should be
given to it.
I put in the equation for a Bouguer disturbance to show that we s

Different approximations for the gravity anomaly can be calculated
depending on the corrections applied to them.
These corrections are usually called \textit{gravity reductions}
and their aim is to approximate the magnitude of gravity at the geoid
$g_{Q^\prime}$ from the magnitude of gravity $g_P$ measured at a point $P$
on or above the surface of the Earth.
The \textit{free-air anomaly}, for example, can be defined as
\citep{blakely1996, hofmann-wellenhof-moritz2005}:

@@

\noindent
where $\partial \gamma/\partial h \approx -0.3086$ mGal/m is the
derivative of the normal gravity with respect to the geometric height $h$
and $H_{P}$ is the orthometric height (i.e., with respect to the geoid)
of point $P$ (Fig. \ref{fig:surfaces}).
The term between parenthesis represents an
approximation of the gravity at the point $Q^{\prime}$ on the geoid.
Similarly, we can approximate the normal gravity $\gamma_{P}$
at a point $P$ on the Earth's surface
from the normal gravity at the ellipsoid $\gamma_{Q}$,
so that the gravity disturbance $\delta g_{P}$
(eq. \ref{eq:gravity-disturbance}) can be rewritten as:

@@

\noindent
where $h_P$ is the geometric height of point P.
Thus, the absolute difference between
the approximated gravity disturbance (eq. \ref{eq:gravity-disturbance-approx}),
at a point $P$ on or close to the Earth's surface,
and the Free-air anomaly (eq. \ref{eq:free-air-anomaly}) is:

@@

The simple Bouguer anomaly is commonly used by geophysicists as the
gravitational effect produced by the gravity sources.
It is defined, over the continents, as follows
\citep{blakely1996, hofmann-wellenhof-moritz2005}:
\begin{equation}
\Delta g_{P}^{B}
= \Delta g_{P}^{F} - 2 \pi G \rho_{t} \, H_{P} \: ,
\label{eq:simple-bouguer-anomaly}
\end{equation}
where $\Delta g_{P}^{F}$ is the Free-air anomaly (eq. \ref{eq:free-air-anomaly})
and the last term on the right side is the simple Bouguer correction.
This term approximates the gravitational attraction that all topographic
masses above the Geoid exerts at the point $P$ by the gravitational
attraction of a homogeneous, infinitely extended slab of
constant density $\rho_{t}$ and thickness equal to $H_{P}$, which is
the orthometric height $H$ (Fig. \ref{fig:surfaces}) at the point $P$.
By computing the simple Bouguer correction at a set of points
of a survey around $P$, we approximately remove,
from the gravity measured at the points of the survey,
the gravitational effect of a homogeneous model representing
the topographic masses above the Geoid.
It is evident from eqs
\ref{eq:free-air-anomaly} to \ref{eq:difference-disturbance-free-air}
that the simple Bouguer anomaly $\Delta g_{P}^{B}$ (eq.
\ref{eq:simple-bouguer-anomaly}) approximates the gravity disturbance
$\delta g_{P}$ (eq. \ref{eq:gravity-disturbance})
minus the gravitational effect produced by this homogeneous topographic
model.
Although this approximation is valid for most practical applications,
it is important to bear in mind not only the terminology
changes, but also the conceptual assumptions.

There was a certain lack of comprehension regarding the
geophysical meaning of gravity anomalies until the
mid 90's.
As properly pointed out by \citet{chapin1996} at that time,
``although the corrections which bring about a Bouguer
gravity anomaly are well established, the reasons for doing
them are not well understood. One cause of this common
misunderstanding is that the subject has been poorly presented in
many of the basic texts".
In his seminal book, \citet{blakely1996} brought some light
on the geophysical meaning of gravity anomalies from the
perspective of applied geophysics. \citet{blakely1996} correctly
defined gravity sources as density contrasts between the actual
internal density distribution of the Earth and the internal density
distribution of the normal Earth.
However, he did not stress that, by removing the normal gravity
evaluated on the ellipsoid from the gravity measured
on the Earth's surface, the remaining anomalous field will reflect
not only the effect produced by the gravity sources, but also a
small combination of gravitational and centrifugal effects.
This additional, non-harmonic and undesired effect is
simply due to the calculation of the normal gravity at a point
other than that were the gravity is measured.

>>>>

\section{The geophysical meaning of gravity reductions}

Different approximations for the gravity anomaly can be calculated
depending on the corrections applied to them.
These corrections are usually called \textit{gravity reductions}
and their aim is to approximate the magnitude of gravity at the geoid
$g_{Q^\prime}$ from the magnitude of gravity $g_P$ measured at a point $P$
on or above the surface of the Earth.

The \textit{free-air anomaly}, for example, can be defined as
\citep{blakely1996, hofmann-wellenhof-moritz2005}:

@@

\noindent
where $\partial \gamma/\partial h \approx -0.3086$ mGal/m is the
derivative of the normal gravity with respect to the geometric height $h$
and $H_{P}$ is the orthometric height (i.e., with respect to the geoid)
of point $P$ (Fig. \ref{fig:surfaces}).
The term between parenthesis represents an
approximation of the gravity at the point $Q^{\prime}$ on the geoid,
disregarding the effect of topographic masses.
Similarly, we can apply the same correction
to calculate the gravity disturbance by
approximating the normal gravity $\gamma_{P}$
at a point $P$ on the Earth's surface
from the normal gravity at the ellipsoid $\gamma_{Q}$:

@@

\noindent
where $h_P$ is the geometric height of point P.
Such an approximation is usually not required because
an analytical expression exists for calculating $\gamma$ at any point
on or above the ellipsoid \citep{li2001}.
The same cannot be said for the free-air anomaly because
there is usually no way of measuring gravity at the geoid ($g_{Q^\prime}$).

Using eqs. \ref{eq:free-air-anomaly} and \ref{eq:gravity-disturbance-approx},
we can estimate the absolute difference between the gravity disturbance and the
free-air anomaly:

@@

The \textit{Bouguer anomaly}
has the aim of removing the gravitational effect of the topographic masses
between the observation point $P$ and the point on the geoid $Q^\prime$.
It is defined in the continents as
\citep{blakely1996, hofmann-wellenhof-moritz2005}:

\begin{equation}
\Delta g^{B} =
\left(g_{P}
    - \frac{\partial \gamma}{\partial h} \, H_{P}
    - 2 \pi G \rho_{t} \, H_{P} \right)
- \gamma_{Q} \: ,
\label{eq:simple-bouguer-anomaly}
\end{equation}

\noindent
where $\rho_t$ is the density, $G$ is the gravitational constant,
and the rightmost term in parenthesis is the simple Bouguer correction.
This term approximates the gravitational effect that all topographic
masses above the geoid exert on point $P$ by the effect of a homogeneous,
infinitely extended slab of constant density $\rho_{t}$ and thickness $H_{P}$.
The same correction can be applied to the gravity disturbance
(using the approximation in eq. \ref{eq:gravity-disturbance-approx})
with the aim of removing the effect of topographic masses between
the observation point $P$ and a point on the ellipsoid $Q$.
We will call this quantity the \textit{Bouguer disturbance}:

\begin{equation}
\delta g_P^{B} \approx
g_{P} -
\left( \gamma_{Q} + \frac{\partial \gamma}{\partial h} h_P \right)
- 2 \pi G \rho_{t} \, h_{P} \: .
\label{eq:bouguer-disturbance}
\end{equation}

At first glance, the only difference between eqs.
\ref{eq:simple-bouguer-anomaly} and \ref{eq:bouguer-disturbance} is the use of
orthometric ($H_P$) versus geometric ($h_P$) heights.
However, their objectives and geophysical interpretations are significantly
different.
The corrections applied when calculating the Bouguer anomaly approximate
gravity at the geoid from measurements made on the Earth's surface by removing
the effects of the distance and the topographic masses between the two points.
Apart from the centrifugal effect mentioned in the previous section,
the gravitational effect of topographic masses between the geoid and ellipsoid
are still present in the anomaly.
Conversely, when calculating the Bouguer disturbance one removes from
the observed gravity the effects of an ellipsoidal reference model (the normal
Earth) and the known topographic masses between the surface of the ellipsoid
and observation point.
Hence, the Bouguer disturbance can be interpreted as being solely the
gravitational effect produced by the density contrasts between the actual
internal density distribution of the Earth and that of the normal Earth.

There was a certain lack of comprehension regarding the geophysical meaning of
gravity anomalies until the mid 90s.
As properly pointed out by \citet{chapin1996},
``although the corrections which bring about a Bouguer gravity anomaly are well
established, the reasons for doing them are not well understood. One cause of
this common misunderstanding is that the subject has been poorly presented in
many of the basic texts''.
In his seminal book, \citet{blakely1996} shed some light on the geophysical
meaning of gravity anomalies from the perspective of applied geophysics and
their relationship to the gravity sources.
We complement this interpretation by stressing the fact that
gravity anomalies reflect not only the gravity sources,
but also a combination of gravitational and centrifugal effects.
These additional, non-harmonic and undesired effects are due to the calculation
of the normal gravity at a point other than that were gravity is measured.
In practice, this issue can be avoided entirely by performing the free-air and
Bouguer corrections using geometric instead of orthometric heights.
